---
title: "TCGA full dataset"
author: "Daniel Skubleny"
date: "05/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data available at cbio portal https://www.cbioportal.org/datasets. Download "Stomach Adenocarcinoma (TCGA, PanCancer Atlas)"

## R Markdown
##Load Data 
```{r, warning= FALSE, message=FALSE}
#Load Data
pheno <- read.delim("~/stad_tcga_pan_can_atlas_2018/data_clinical_patient.txt", comment.char="#")
edata <- read.csv("~/stad_tcga_pan_can_atlas_2018/data_RNA_Seq_v2_expression_median_ENTREZUPDATED.csv")
pathology <- read.delim("~/stad_tcga_pan_can_atlas_2018/data_clinical_sample.txt", comment.char="#")
```
```{r}
#Make feature data table for future reference and annotation 
fdata <- read.csv("~/stad_tcga_pan_can_atlas_2018/data_RNA_Seq_v2_expression_median_ENTREZUPDATED.csv")
fdata = fdata %>% dplyr::select(Hugo_Symbol,Entrez_Gene_Id)
```


```{r}
#Annotate entrez ID
annotatedfdata = AnnotationDbi::select(org.Hs.eg.db,keys = as.character(edata$Entrez_Gene_Id), columns = c("SYMBOL","ENTREZID","ENSEMBL"), keytype = "ENTREZID")
```
```{r}
# Change column name to ENTREZID in edata 
colnames(edata)[which(names(edata) == "Entrez_Gene_Id")] <- "ENTREZID"

edata = merge(annotatedfdata,edata, by.x="ENTREZID")
```

```{r}
#Remove EnsembleID, HUGO and Entrez ID
edata = dplyr::select(edata, -c("Hugo_Symbol","ENTREZID","ENSEMBL")) # Remove HUGO columns
```
## Check for NA
```{r}
table(is.na(edata$SYMBOL))
# There is NA due to Entrez Ids that no longer map to gene symbols. 
```
```{r}
#Remove NA Symbols
edata = edata[!is.na(edata$SYMBOL),]
```
## Check for duplicates
```{r}
table(duplicated(edata$SYMBOL))
```
```{r}
#Aggregate duplicate genes by mean
edata %>% group_by(SYMBOL) %>% summarise_all(mean) %>% data.frame() -> edata_merged
edata = edata_merged
 #Aggregate and mean duplicate ENTREZIDs 
```



##Check for NA genes and NA feature data
```{r}
table(is.na(edata$SYMBOL))
#There is NA Entrez Genes 
```

```{r}
table(is.na(edata))
#There is NA data. We will need to investigate this later on. 
```

## Make gene SYMBOL row names
```{r}
#This ensures we can set up a matrix for future procedures
edata = data.frame(edata, row.names = "SYMBOL")
```
## Assess dimensions
```{r}
#Assess dimensions 
dim(edata)
```

```{r}
edata_fullTCGA = edata
```

Boxplot
```{r}
# Look at data that is not transformed vs transformed
par(mfrow=c(1,2))
boxplot(edata_fullTCGA, col=2, range=0)
boxplot(log2(edata_fullTCGA+1), col=2, range=0)
```
Histogram
```{r}
par(mfrow=c(1,2))
hist(edata_fullTCGA[,1],col=2)
hist(log2(edata_fullTCGA[,1]+1),col=3)
```
Density Plot
```{r}
par(mfrow=c(1,2))
omitted = na.omit(edata_fullTCGA)
plot(density(omitted[,1]),col=2)
lines(density(omitted[,2]),col=3)
plot(density(log2(omitted[,1]+1)),col=2)
lines(density(log2(omitted[,2]+1)),col=3)
```
qqPlot
```{r}
par(mfrow=c(1,2))
qqplot(edata_fullTCGA[,1], edata_fullTCGA[,2],col=3)
qqplot(log2(edata_fullTCGA[,1]+1), log2(edata_fullTCGA[,2]+1),col=3)
```
Bland Altman Plot
```{r}
# Bland Altman Plot
mm = edata_fullTCGA[,1] - edata_fullTCGA[,2]
aa = edata_fullTCGA[,1] + edata_fullTCGA[,2]
plot(aa,mm,col=2) 
```
Let's reassess after a log2 transform 
```{r}
# Bland Altman Plot
mm = log2(edata_fullTCGA[,1]+1) - log2(edata_fullTCGA[,2]+1)
aa = log2(edata_fullTCGA[,1]+1) + log2(edata_fullTCGA[,2]+1)
plot(aa,mm,col=2) 
```

##Transform data by log2 based on exploratory data analysis. 
```{r}
log_edata_fullTCGA = log2(edata_fullTCGA+1)
```

## Here we will remove low abundance genes because we know that concordance between nanostring and microarray or rnaseq platforms is best with higher abudnant, high dynamic range and more varaiable genes (Kwing et al. Biological Validation of RNA Sequencing Data From Formalin-Fixed Paraffin-Embedded Primary Melanomas, Richard et al. Comparison of gene expression microarray data with count-based RNA measurements informs microarray interpretation, Illumina. RNA-Seq Offers Significant Advantages in Immuno-Oncology Research)

#Median intensity 
```{r}
median_fullTCGA = rowMedians(as.matrix(log_edata_fullTCGA))
hist(median_fullTCGA, breaks=200)
#Given this is RNAseq count data I will filter low expression genes by mean. 
```


```{r}
log_edata_fullTCGA = as.data.frame(log_edata_fullTCGA)
log_edata_fullTCGA = filter(log_edata_fullTCGA,rowMeans(log_edata_fullTCGA) > 1)
dim(log_edata_fullTCGA)
```
```{r}
log_edata_fullTCGA = log_edata_fullTCGA[which(rowMeans(!is.na(log_edata_fullTCGA)) > 0.2), which(colMeans(!is.na(log_edata_fullTCGA)) > 0.2)]
dim(log_edata_fullTCGA)
#No significant missing data 
```
```{r}
table(is.na(log_edata_fullTCGA))
#No missing data. Do not need imputation 
```

```{r}
median_fullTCGA2 = rowMedians(as.matrix(log_edata_fullTCGA))
hist(median_fullTCGA2, breaks=200)
#New distribution after removal of rowmean < 1
```


From this we determine that there are no higher abundance genes that require imputation. There are no data missingness. There are no NA. 


#Quantile Normalization 

```{r}
#Check if quantile normalization is required.
colramp = colorRampPalette(c(3,"white",2))(20)
plot(density(log_edata_fullTCGA[,1]),col=colramp[1],lwd=1,ylim=c(0,.20))
for(i in 2:300){lines(density(log_edata_fullTCGA[,i]),lwd=1,col=colramp[i])}
```



